Exploring dimensions of biodiversity in Japanese ferns

### 2021-08-26 ms/manuscript.Rmd
Update MS
per comments from BM and AE 2021-08-26

<!-- may include title words -->
**Key words:** apomixis, biodiversity, biogeography, ferns, Japan, phylogenetic diversity.

@@

Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology with two equally important aspects: to understand the processes generating it, and to conserve it. 
The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on species richness. 
Hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 34 terrestrial hotspots [@Myers2000] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006]. 
However, richness alone cannot provide a complete picture of biodiversity. 
All organisms are related to a greater or lesser degree by descent from a common ancestor, and these evolutionary relationships must be taken into account to obtain a full understanding of the biodiversity present in an area.

@@

Our data include all ferns known to occur within the boundaries of Japan (20°N to 45°N, 122°E to 153°E). 
Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). 
The main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ryukyu and Ogasawara archipelagos are oceanic islands that have never been connected to the mainland. 
The climate is subtropical in the south and temperate in other parts of the country. 
Map data were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).

@@

We used a list of all non-hybrid fern specimens housed at herbaria in Japan. 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated). 
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).

@@

We measured phylogenetic diversity (PD) using the Faith's PD (the total branch length connecting all terminal taxa in a community, including the root of the phylogeny) [@Faith1992]. 
To detect areas with unusually long or short branch lengths, we calculated relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].

@@

We measured phylogenetic endemism (PE) as the sum of all branches at a site inversely weighted by their range [@Rosauer2009].

Observed values of PD, RPD, FD, RFD, and PE are known to be closely related to species richness: as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others. 
Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of taxa by conducting a randomization test. 
We generated 999 random communities using the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values. 
Statistical significance was assessed using a one-tailed test for PE or a two-tailed test for other metrics with alpha = 0.5; observed values in the extreme 5% of the null distribution were considered significant. 
We also calculated the standard effect size (SES) of each biodiversity metric. 
SES measures how extreme the observed value (*obs*) is relative to the null distribution (*null*):

@@

We tested whether endemic areas have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleo-endemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014]. 
Low RPE indicates short branchlengths (neoendemism) and high RPE indicates long branchlengths (paleoendemism). 
Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable relative to those calculated for the null distribution.

@@

It is possible that endemism may be affected by species that are actually wide-ranging, but only have a small range in Japan (a "border effect"). 
This is especially expected for southern, sub-tropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country. 
To partially account for this, we also calculated all endemism-related metrics using a dataset including only species endemic to Japan.

@@

We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html). 
We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" (some human activities allowed by permit). 
We cropped grid-cells to only include land area, then calculated the percent of protected area of each status type within grid-cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness. 
For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above. 
For taxonomic richness, cells ranked in the top >5% were considered highly diverse.

@@

# Caculate number of native, non-hybrid fern taxa
total_fern_taxa <- fern_list %>%
  filter(hybrid == FALSE) %>%
  pull(taxon) %>%
  n_distinct()

@@

# Calculate number/percent of missing data in trait matrix
fern_traits_missing_summary <-
  fern_traits %>%
  pivot_longer(-taxon) %>%
  mutate(missing_data = is.na(value)) %>%
  summarize(
    total = n(),
    missing = sum(missing_data)
  )

@@

The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent`) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families of native Japanese ferns. 
The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* (families follow @PteridophytePhylogenyGroupI2016). 
The age of the fern clade is estimated to be `r phylogeny_age` ma.

@@

After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits) (`r table("traits-used")`). 
The trait matrix included `r nrow(fern_traits)` taxa and `r fern_traits %>% select(-taxon) %>% ncol()` traits, with `r percent_traits_missing` missing data.

@@

Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`). 
The vast majority of grid-cells with significant endemism are mixed.
Concentrations of paleoendemism were observed in the Ryukyu archipelago, on Okinawa, Miyako, and Iriomote Islands. 
Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in Hokkaido (`r figure("endemism")`). 
When we ran CANAPE with a reduced dataset including only taxa absolutely endemic to Japan to account for the border effect, there were fewer significant cells, but a similar pattern of significantly endemic cells mostly occurring in the southern islands was observed (`r s_figure("endemism-restricted")`).

@@

Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`b). 
The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a). The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).

Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (`r figure("modelfit")`g).
Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in southern Honshu (`r s_figure("genus-map")`).
This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns. 
We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), consistent with an effect on genetic diversity (`r s_figure("modelfit-not-ult")`b).
While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`). 
Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca. 34.8 °N latitude and increases with temperature in Japan (their study was limited to Honshu and did not include the southern islands). 
@Tanaka2014 asserted that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses.

@@

The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`). 
Interestingly, the bioregions identified here correspond roughly to major forest types previously categorized on an *ad-hoc* basis; Cluster 1 corresponds to the cool or subartic deciduous type, Cluster 2 to the warm evergreen type, and Clusters 3 and 4 to the subtropical rainy type [@Shimizu2014]. 
This suggests that ferns are useful as possible bioindicators in Japan. 
Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions. 
The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits clusters 2 and 3, there is some overlap between them on the Amami-Oshima Isl. (`r figure("bioregions")`). 
The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic Cluster 3 (mostly the Ryukyu islands; wetter) from Cluster 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).

@@

Although the different facets of biodiversity have different distribution patterns, we find that grid-cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric (`r figure("conserv-status")`). 
One possible reason is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas (A. Ebihara, pers. comm.). 
Another (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.

Despite the good news that ferns enjoy relatively strong protection status in Japan, we must bear in mind that there are other threats posed that conservation zones cannot ameliorate, the most obvious being climate change. 
One such other threat is herbivory by native Japanese deer, *Cervus nippon*. 
Populations of these deer have increased rapidly since the 1970s due to a combination of climate change and a decrease in human hunters and natural predators [@Takatsuki2009]. 
At dense populations, the deer can prevent forest regeneration and result in plant communities dominated by a limited number of species unpalatable to deer [@Takatsuki2009], although lower population sizes may promote plant diversity as per the intermediate disturbance hypothesis [@Nishizawa2016].
They have caused extensive damage even within national parks [@Tokita2006], which have the highest protection status. 
Although various efforts to prevent deer herbivory have been deployed such as culling and exclusion fences [@Takatsuki2009], much work remains to be done to save vulnerable plant populations.

@@

There are a number of caveats that must be kept in mind when interpreting these results, but these also suggest avenues for future research. 
First, the data include species collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan. 
Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (`r s_figure("abundance")`). 
Indeed, one promising avenue of research may be to take advantage of the most densely sampled areas to investigate changes in abundance over time.

@@

One piece of biodiversity that we did not consider here is hybrids. 
The Japanese fern flora includes `r n_hybrid_fern_taxa` hybrid taxa. 
Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification. 
Future studies should consider the distribution of hybrids in Japan and their possible effects on biodiversity.
Methods for calculating PD would need to be modified to take into account multiple lineages contributing to a terminal taxon.

@@

**`r figure("raw-diversity")`**. Raw taxonomic (a), phylogenetic (b), and functional (c) diversity of the ferns of Japan. For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.

@@

**`r figure("rand-diversity")`**. Results of randomization test for phylogenetic and functional diversity of the ferns of Japan. (a), Phylogenetic diversity (PD). (b), Relative phylogenetic diveristy (RPD). (c), Functional diveristy (FD). (d), relative functional diveristy (RFD). For each metric, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test. For details, see Materials and Methods.

@@

**`r figure("endemism")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism). 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches. For details, see Materials and Methods.

@@

**`r figure("conserv-status")`**. Percent of land area with protected status for grid cells with significantly high biodiversity. Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE). Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Methods). Vertical lines indicate total percent area protected across Japan regardless of degree of biodiversity; dashed, high protection level (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`); dotted, medium protection level (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`).

>>>>

<!-- may include title words -->
**Key words:** apomixis, biodiversity, biogeography, CANAPE, ferns, Japan, phylogenetic diversity.

@@

Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology with two equally important aspects: to understand the processes generating it, and to conserve it. 
The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on species richness. 
Hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 36 terrestrial hotspots [@Myers2000; @Noss2016] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006]. 
However, richness alone cannot provide a complete picture of biodiversity. 
All organisms are related to a greater or lesser degree by descent from a common ancestor, and these evolutionary relationships must be taken into account to obtain a full understanding of the biodiversity present in an area.

@@

Our data include all ferns known to occur within the boundaries of Japan (20°N to 45°N, 122°E to 153°E). 
Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). 
Most of the main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ogasawara archipelago consists of oceanic islands that have never been connected to the mainland. 
The climate is subtropical in the south and temperate in other parts of the country. 
Map data were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).

@@

We used a list of all non-hybrid fern specimens deposited at TNS (herbarium codes follow Thiers2021). 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated). 
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).

@@

We measured phylogenetic diversity (PD) using the Faith's PD [the total branch length connecting all terminal taxa in a community, including the root of the phylogeny; @Faith1992]. 
To detect areas with unusually long or short branches, we calculated relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].

@@

We measured phylogenetic endemism (PE) as the sum of all branches at a site inversely weighted by their range size [@Rosauer2009].

Observed values of PD, RPD, FD, RFD, and PE are known to be related to species richness: as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others. 
Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of taxa by conducting a randomization test. 
We generated 999 random communities using the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values. 
Statistical significance was assessed using a one-tailed test for PE or a two-tailed test for other metrics with alpha = 0.5; observed values in the extreme 5% of the null distribution were considered significant. 
We also calculated the standard effect size (SES) of each biodiversity metric. 
SES measures how extreme the observed value (*obs*) is relative to the null distribution (*null*):

@@

We tested whether centers of endemism have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleo-endemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014]. 
Low RPE indicates short range-restricted branches (neoendemism) and high RPE indicates long range-restricted branches (paleoendemism). 
Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable relative to those calculated for the null distribution.

@@

It is possible that endemism may be affected by species that are actually wide-ranging, but only have a small range in Japan (a "border effect"). 
This is especially expected for southern, sub-tropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country. 
To partially account for this, we also calculated all endemism-related metrics using a dataset including species found only in Japan.

@@

We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html). 
We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" (some economic activities allowed by permit) following [@Kusumoto2017]. 
We cropped grid-cells to only include land area, then calculated the percent of protected area of each status type within grid-cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness. 
For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above. 
For taxonomic richness, cells ranked in the top >5% were considered highly diverse.

@@

fern_list_non_hybrid <-
  fern_list %>%
  filter(hybrid == FALSE) %>%
  mutate(gn_parse_tidy(.$scientific_name) %>% select(taxon = canonicalsimple)) %>%
  mutate(taxon = str_replace_all(taxon, " ", "_"))

# Caculate number of native, non-hybrid fern taxa
total_fern_taxa <- fern_list_non_hybrid %>%
  pull(taxon) %>%
  n_distinct()

@@

# Calculate number/percent of missing data in trait matrix
trait_sampling_summary <-
fern_list_non_hybrid %>%
  select(taxon) %>%
  verify(all(fern_traits$taxon %in% taxon)) %>%
  left_join(
    transmute(fern_traits, taxon, sampled = TRUE), by = "taxon"
  ) %>%
  mutate(sampled = replace_na(sampled, FALSE))

n_taxa_traits_sampled <- sum(trait_sampling_summary$sampled)
percent_taxa_traits_sampled <- (n_taxa_traits_sampled / total_fern_taxa) %>% scales::percent(accuracy = 0.1)

fern_traits_missing_summary <-
  fern_traits %>%
  pivot_longer(-taxon) %>%
  mutate(missing_data = is.na(value)) %>%
  summarize(
    total = n(),
    missing = sum(missing_data)
  )

@@

The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families. 
The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* (families follow @PteridophytePhylogenyGroupI2016). 
The age of the fern clade is estimated to be `r phylogeny_age` ma.

@@

After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits) (`r table("traits-used")`). 
The trait matrix included `r n_taxa_traits_sampled` taxa (`r percent_taxa_traits_sampled`) and `r fern_traits %>% select(-taxon) %>% ncol()` traits, with `r percent_traits_missing` missing data.

@@

Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`). 
The vast majority of grid-cells with significant endemism are mixed.
Concentrations of paleoendemism were observed in the Ryukyu archipelago, on Okinawa, Miyako, and Iriomote Islands. 
Small clusters of cells containing significant endemism were also observed northwest of Tokyo (Nagano prefecture), and in Hokkaido (`r figure("endemism")`). 
When we ran CANAPE with a reduced dataset including only taxa absolutely restricted to Japan to account for the border effect, there were fewer significant cells, but a similar pattern of significant endemism mostly occurring in the southern islands was observed (`r s_figure("endemism-restricted")`).

@@

Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half of Honshu, and particularly on Kyushu (`r figure("rand-diversity")`b). 
The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a). The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).

Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (`r figure("modelfit")`g), and to a slightly lesser extent SES of PD (`r figure("modelfit")`f).
Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in southern Honshu (`r s_figure("genus-map")`).
This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns. 
We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (`r s_figure("modelfit-not-ult")`b).
While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`). 
Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca. 34.8 °N latitude and increases with temperature in Japan (their study was limited to Honshu and did not include the southern islands). 
@Tanaka2014 asserted that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses.

@@

The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`). 
Interestingly, the bioregions identified here correspond roughly to major forest types previously categorized on an *ad-hoc* basis; Cluster 1 corresponds to the cool or subartic deciduous type, Cluster 2 to the warm evergreen type, and Clusters 3 and 4 to the subtropical rainy type [@Shimizu2014]. 
This suggests that ferns are useful as bioindicators in Japan. 
Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions. 
The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits clusters 2 and 3, there is some overlap between them on the Amami-Oshima Isl. (`r figure("bioregions")`). 
The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic Cluster 3 (mostly the Ryukyu islands; wetter) from Cluster 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).

@@

Although the different facets of biodiversity have different distribution patterns, we found that grid-cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric (`r figure("conserv-status")`). 
One possible reason is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas (A. Ebihara, pers. comm.). 
Another (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.

Despite the good news that ferns enjoy relatively strong protection status in Japan, we must bear in mind that there are other threats posed that conservation zones cannot ameliorate, the most obvious being climate change. 
One other such threat is herbivory by native Japanese deer, *Cervus nippon*. 
Populations of these deer have increased rapidly since the 1970s due to a combination of climate change and a decrease in human hunters and natural predators [@Takatsuki2009]. 
At dense populations, the deer can prevent forest regeneration and result in plant communities dominated by a limited number of species unpalatable to deer [@Takatsuki2009], although lower population sizes may promote plant diversity as per the intermediate disturbance hypothesis [@Nishizawa2016].
They have caused extensive damage even within national parks [@Tokita2006], which have the highest protection status. 
Although various efforts to prevent deer herbivory have been deployed such as culling and exclusion fences [@Takatsuki2009], much work remains to be done to save vulnerable plant populations.

@@

There are a number of caveats that must be kept in mind when interpreting these results, but these also suggest avenues for future research. 
First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan. 
Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (`r s_figure("abundance")`). 
Indeed, one promising avenue of research may be to take advantage of the most densely sampled areas to investigate changes in abundance over time.

@@

One element of biodiversity that we did not consider here is hybrids. 
The Japanese fern flora includes `r n_hybrid_fern_taxa` hybrid taxa. 
Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989]. 
Future studies should consider the distribution of hybrids in Japan and their possible effects on biodiversity.
Methods for calculating PD would need to be modified to take into account multiple lineages contributing to one terminal taxon.

@@

**`r figure("raw-diversity")`**. (a) Raw taxonomic diversity (b), phylogenetic diversity, and (c) functional diversity of the ferns of Japan. For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.

@@

**`r figure("rand-diversity")`**. Results of randomization test for phylogenetic and functional diversity of the ferns of Japan. (a), Phylogenetic diversity (PD). (b), Relative phylogenetic diversity (RPD). (c), Functional diversity (FD). (d), relative functional diversity (RFD). For each metric, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test. For details, see Materials and Methods.

@@

**`r figure("endemism")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism). 'paleo' indicates areas with an overabundance of rare long branches; 'neo' indicates areas with an overabundance of rare short branches; 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches. For details, see Materials and Methods.

@@

**`r figure("conserv-status")`**. Percent of land area with protected status for grid cells with significantly high biodiversity by at least one metric. Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE). Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Materials and Methods). Vertical lines indicate total percent area protected across Japan regardless of degree of biodiversity; dashed, high protection level (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`); dotted, medium protection level (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`).
